---
title: "MA Public School Project"
output: html_document
date: '2022-11-22'
---


## Concept

Im seeking to answer the following question: Does ethnicity make up
of a school district effect art courses offered, sat performance, disciplinary 
action, and teacher salaries?

In this document I will:

1. Load data from the [Massachusetts School and District Data](https://www.doe.mass.edu/SchDistrictData.html)
2. Process different variables into one data set
3. And make initial visualizations

## Required Packages

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
```


## Data Manipulation
Load in ethnicity data and add a column for African American & Hispanic.
```{r}
ClassSizebyEthnicity <- read.csv("raw_data/ClassSizebyRaceEthnicity.csv")
```

```{r, cache=TRUE}
district_race <- ClassSizebyEthnicity %>%
  select(District.Name, African.American.., Hispanic.., White..) %>%
  group_by(District.Name) %>%
  mutate(percent_black_hispanic = sum(African.American.., Hispanic.., na.rm = TRUE ))
```

Read in data for art courses taken and add a column for percent of students that take art classes.
```{r cache=TRUE}
artcourse <- read.csv("raw_data/artcourse.csv")
```


```{r, cache=TRUE}
art_course_taken <- artcourse %>%
  select(District.Name, All.Grades, Total.Students) %>%
  mutate(percent_students_in_art = (All.Grades / Total.Students)*100)
```

Read in sat performance data and join it with art course data, then join that data set to ethnicity data set.
```{r cache=TRUE}
sat_performance <- read.csv("raw_data/sat_performance.csv")
```

```{r cache=TRUE}
 sat_art <- select(sat_performance, Tests.Taken, District.Name) %>%
  inner_join(art_course_taken, by = "District.Name")
```

```{r cache=TRUE}
 art_sat_race <- inner_join(sat_art,district_race, by = "District.Name") %>%
   select(-African.American.., -Hispanic..)
```

Read in student discipline data and join it to the above data set
```{r, cache= TRUE}
cheese <- read.csv("raw_data/StudentDisciplineDataReport.csv") %>%
  select(District.Name, Students.Disciplined) %>%
  inner_join(art_sat_race, by = "District.Name")
```

Read in teacher salary data and join it to the above data set
```{r, cache= TRUE}
grouped_data <- read.csv("raw_data/TeacherSalaries.csv") %>%
  select(District.Name,Average.Salary) %>%
  inner_join(cheese, by = "District.Name")
```

Change sat and student discipline data to proportions.
```{r, cache= TRUE}
proportion_data <- mutate(grouped_data, percent_take_sat = (Tests.Taken/Total.Students)*100) %>%
  mutate(percent_disciplined = (Students.Disciplined / Total.Students)*100)%>%
  select(-Students.Disciplined, -Tests.Taken, -All.Grades)
```

## Data Plots
```{r,echo=FALSE}
knitr::kable(
  proportion_data[1:5, ], 
  caption = "proportion data set"
)

```{r, echo= FALSE, message=FALSE}
ggplot(grouped_data, aes(percent_black_hispanic))+
  geom_histogram()
```